---
title: "Ch. 5 Data Transformation"
author: "Jerri Schorr"
date: 09-31-2019
output: github_document
---

```{r setup chunk, include=FALSE}
library(tidyverse)
library(nycflights13)
```


##### **Chapter 5**

###### **5.1**

* It is rare to get data exactly as you need it. 
    + we will learn how to reorder the data and create new variables or summaries of the data
    + we will use stats::filter() & stats::lag()
    
```{r}
nycflights13::flights
```

* This data set is a **tibble** 
    + it is a little different so that it works well with tidyverse
    
* 5 dplyr functions
    + filter()
    + arrange()
    + select()
    + mutate()
    + summarise()
    
###### **5.2** 

* Arguments and notation
    + first argument is data frame
    + all other arguments are what rows we want to filter out
    + Example:
    
```{r}
jan1 <- filter(flights, month == 1, day == 1)
```
    + we get all the data from january 1st
    + we assigned this filtered data frame to jan1
    
* Exercises 

```{r}
filter(flights, arr_delay >= 2)
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, carrier == "UA" | carrier == "AA" | carrier == "DL")
filter(flights, month == 7 | month == 8 | month == 9)


```

###### **5.3**

* arrange is similar to filter but it changes order of rows instead of selecting them 

* Exercises

```{r}
arrange(flights, is.na(month))
arrange(flights, dep_delay)
arrange(flights, desc(dep_delay))

```

###### **5.4**

* select() will select columns

###### **5.5**

* add new variables with the muatate function

    + you can refer to columns that you have just created
    + transmute lets you keep new variable only
    
###### **5.6**

* summarise makes the data frame a single row

* we can use the **pipe** operator to combine multiple functions at once
    + **%>%**
    
###### **5.7**

* we can use the group function with mutate() and filter() as weel as the summarise() function



